# Reproducible Research: Peer Assessment 1
by Frederik Sua

## Loading and preprocessing the data

```{r}
url <- "http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(url, "repdata-data-activity.zip", mode="wb")
unzip("repdata-data-activity.zip")
activity <- read.csv("activity.csv")
```

## What is the mean number of steps taken per day?

1. Make a histogram of steps per day.

```{r}
steps.date <- aggregate(steps ~ date, data = activity, FUN = sum)
par(las=2) # make label text perpendicular to axis
barplot(steps.date$steps, main="Steps Per Day", names.arg = steps.date$date, ylab = "steps")
```

2. Calculate the mean and median steps per day.

```{r}
mean(steps.date$steps)
median(steps.date$steps)
```

## What is the average daily activity pattern?

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
steps.time <- aggregate(steps ~ interval, data = activity, FUN = mean)
plot(steps.time, type = "l")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
steps.time$time[which.max(steps.time$steps)]
```
## Inputing missing values

Number of rows with NAs

```{r}
sum(is.na(activity))
```

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

Using 5 min intervals to fill in missing values

```{r}
activity <- merge(activity, steps.time, by = "interval", suffixes = c("", 
    ".y"))
nas <- is.na(activity$steps)
activity$steps[nas] <- activity$steps.y[nas]
activity <- activity[, c(1:3)]
```
Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}
steps.date <- aggregate(steps ~ date, data = activity, FUN = sum)
par(las=2) # make label text perpendicular to axis
barplot(steps.date$steps, main="Steps Per Day", names.arg = steps.date$date, ylab = "steps")
```

Output mean and median

```{r}
mean(steps.date$steps)
median(steps.date$steps)
```

The mean and median of the step counts seem to be comparable with or without filling in these values, the impact appears to be insignificant.

## Are there differences in activity patterns between weekdays and weekends?

Create a factor variable definine a day as a weekday or a weekend day

```{r}
daytype <- function(date) {
    if (weekdays(as.Date(date)) %in% c("Saturday", "Sunday")) {
        "weekend"
    } else {
        "weekday"
    }
}
activity$daytype <- as.factor(sapply(activity$date, daytype))
```

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
par(mfrow = c(2, 1))
for (type in c("weekend", "weekday")) {
    steps.type <- aggregate(steps ~ interval, data = activity, subset = activity$daytype == 
        type, FUN = mean)
    plot(steps.type, type = "l", main = type)
}
```

Activity seems to peak on weekends and stay constant, while on weekdays there's normally only a single period of higher levels of activity.
